Algebraic Surfaces

### 2016-05-22 project/article_revised_revised/main.tex
fixed some of the comments

\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx} 
\makeindex

@@

\usepackage[]{cleveref}

@@

    In algebraic geometry the central objects of study is the set of solutions
    to polynomial equations. Finding the zeros of, as an example, the quadratic
    equation
    \begin{equation}
        \notag
        ax^2 + bx + c = 0
    \end{equation}
    is something most people that have been exposed to elementary mathematics
    have done. Later on, one learns how to find the zeroes of more general
    functions, of either one or several variables over some general field of
    coefficients. This leads us to our first definition.

    \begin{definition}[Polynomial ring\index{polynomial ring} \cite{Pol16}]
        A \emph{polynomial ring} in $n$ variables over a field $K$ is the set
        of all functions on the form
        \begin{equation}
            \notag
            f(x_1, \ldots, x_n) = \sum^{}_{\alpha} p_{\alpha}\prod_{i=1}^{n} x_n^{\alpha_i}
        \end{equation}
        where $\alpha = (\alpha_1, \alpha_2, \ldots, \alpha_n)$ is a
        multi-index, and $p_\alpha = p_{\alpha_1,\ldots,\alpha_n}$ is an
        element in $K$.  We denote this polynomial ring $K[x_1, x_2, \ldots,
        x_n]$.
    \end{definition} 
    As an example, the most familiar polynomial ring is $\R[x]$. Elements of
    this ring include
    \begin{align*}
        f_1(x) = x^2, && f_2(x) = \pi, && f_3(x) = x^{17} + x^{3} - e^{4}, && f_4(x) = x^2 + 1.
    \end{align*} 
    Note that some of these have zeroes in the real numbers $\R$ while some
    does not. Most notably is the function $f_4$ whose zeroes lie in the realm
    of complex numbers, $\C$.

    Given a set of polynomials, $f_1, f_2, \ldots, f_m$ in a polynomial ring
    $K[x_1, x_2, \ldots, x_n]$ we can look at their common zeroes. It is in our
    interest to restrict our attention to polynomial rings where the field of
    coefficients is algebraically closed. We recall the definition:

    \begin{definition}[Algebraically closed field {\cite[p.~287]{FK03}}\index{algebraically closed field}]
        A field $K$ is \emph{algebraically closed} if every non-constant
        polynomial in $K[x_1, x_2, \ldots, x_n]$ has a zero in $K$.
    \end{definition}
    Given an algebraically closed field $K$, we define \emph{affine
    $n$-space}\index{affine $n$-space} over $K$ to be the set of all
    $n$-tuples of elements of $K$. We denote affine $n$-space by $\A_K^n$ or
    $\A^n$ if the field of coefficients is clear from context.  For our intents
    and purposes, we will be working over the algebraically closed field of
    complex numbers $\C$.

@@

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities.
    \begin{definition}[Singularity {\cite[p.~3]{Whi08}}\index{singularity}]
        A singularity of the variety $V = \V(f_1, f_2, \ldots, f_m) \subset
        \A_K^n$ is a point $p \in V$ such that the Jacobian matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f_1}{\partial x_1} & \hdots & \frac{\partial f_1}{\partial x_n} \\
                \vdots & \ddots & \vdots \\
                \frac{\partial f_m}{\partial x_1} & \hdots & \frac{\partial f_m}{\partial x_n}
            \end{bmatrix}(p)
        \end{equation}
        has rank strictly less than $\min(m, n)$.
    \end{definition}

    A very useful consequence of this definition deal with the varieties of
    single polynomials and will aid us tremendously in the search for
    singularities.
    \begin{corollary}
        \label{crl:single_poly}
        A point $p$ is a singularity of the variety $V = \V(f) \subset \A_K^n$
        if and only if the partial derivatives of $f$ vanish identically at
        $p$. 
    \end{corollary}
    \begin{proof}
        We recall that the rank of a matrix is the dimension of its column
        space, that is for a matrix $A$, $\rank(A) = \dim(\Col(A))$.  Assume
        first that $p$ is a singularity of $V$. Then $\rank(A) < min(n, m) =
        1$. This means that the matrix $A$ has no independent columns,
        otherwise the columns of $A$ would span a space of dimension greater
        than or equal to one. The Jacobian matrix reads in this case
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f}{\partial x_1}(p) & \hdots & \frac{\partial f}{\partial x_n}(p)
            \end{bmatrix}.
        \end{equation}
        Since the matrix has no independent columns, we cannot have any pivot
        columns, hence each entry must be zero. Consequently, all partial
        derivatives of $f$ vanish at $p$.  For the opposite implication, assume
        that the partials vanish identically at $p$. Then the Jacobian matrix
        reads
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                0 & \hdots & 0
            \end{bmatrix},
        \end{equation} and consequently $A$ has rank strictly less than one.
        Hence $p$ is a point of singularity of $V$.
    \end{proof}

@@

    \section{Algebraic Surfaces}
    \label{sec:algebraic_surfaces}

@@

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ''The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ''The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{Curves not corresponding to functions.}
        \label{fig:algebraic_curves}
    \end{figure}

@@

    From \cite[p.~7]{Lab14} we have the following relation on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(d) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

@@

    We call a variety that exhibits no singularities \emph{smooth}\index{smooth
    variety}. It is often possible to study a smooth copy of a variety with
    singularities if we move from affine space to projective space. This
    process can be defined in many different ways, with the seemingly most
    prevalent one being through maps from various spaces to each other. In this
    paper we will try to keep it as intuitive as possible, by using the more
    geometrically based procedure. We first define the notion of a projective
    space.

    \subsection{Projective space}
    \label{sub:projective_space}

    Informally, projective space generalizes the notion of parallel lines
    intersecting at infinity. The general definition of projective $n$-space is
    the set of lines in $\A^{n+1}$ passing through the origin. For our purposes
    we will take $\A^{n+1} = \C^{n+1}$.

@@

    We start by examining resolutions of single points. Informally, when we
    blow up a point $p$ in $\A^n$, we replace the point $p$ by an entire copy
    of a projective space $\P^{n-1}$ and try to leave the rest of the space
    $\A^n$ ''unchanged''. We will give a few examples, that hopefully will
    illustrate the geometrical aspect of this, but first a formal definition.

    \begin{definition}[Blowup of $\A^{n}$ at a point $p$.]
        We let the \emph{blowup-surface}\index{blowup-surface!in a point}
        $\Bl_p\A^{n}$ be the set of all pairs consisting of a point $p \in
        \A^n$ and a line from $p$ through the origin. The surface is embedded
        in the space $\A^n\times\P^{n-1}$. We write
        \begin{equation}
            \notag
            \Bl_p\A^n = \left\{ (p, L) \mid p \in \A^{n} \text{ and } p \in L \right\}.
        \end{equation}
    \end{definition}
    Now, this is a fairly technical definition, so where do we go from here?
    When blowing up a surface at a point $p$ we are essentially giving the
    surface ''more room'' so that the points creating the singularities can
    pass each other without hassle. In order to aid us with the above
    definition, we show the following claim that tells us under what
    circumstances the point $p$ lie on the line $L$:

    \begin{proposition}
        \label{prop:point_line}
        Given a point $x = (x_1, x_2, \ldots, x_n)$ in $\A^n$ and a point $y =
        [y_1 : y_2 : \ldots : y_n]$ in the space $\P^{n-1}$, the point $x$ is a
        multiple of $y$ if and only if the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x_1 & x_2 & \hdots & x_n \\
                y_1 & y_2 & \hdots & y_n
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to one. Similarly, the rank is less than or
        equal to one only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
        $1 \leq i < j \leq n$.
    \end{proposition} 
    \begin{proof}
        In order to verify this, we recall that $\rank(A) = \rank(A^T)$, so we
        chose to work with the transposed matrix instead. Assume that $x$ is a
        multiple of $y$, so $y = \lambda x$ for some scalar $\lambda$. Then
        \begin{equation}
            \notag
            A^T = \begin{bmatrix}
                x_1 \;\, \lambda x_1 \\
                x_2 \;\, \lambda x_2 \\
                \vdots \\
                x_n \;\, \lambda x_n
            \end{bmatrix}
        \end{equation}
        is the associated matrix. If $x = 0$, then $A^T$ is the zero matrix and
        has therefore rank 0. If $x \neq 0$, then we have exactly one pivot
        column in this matrix, hence $\dim(\Col(A^T)) = 1$. In any case,
        $\rank(A) = \rank(A^T) \leq 1$. For the opposite implication. If $A$
        has rank 0, then $A$ is the zero matrix, and hence trivially, $x$ is a
        multiple of $y$, both being zero. If $A$ has rank $1$, then there is
        one linearly independent column in $A^T$, hence the other column can be
        expressed as a multiple of the first. In other words, $x$ is a multiple
        of $y$.

@@

        Setting $u = 1$ we get the equation $y = xv$. This can be substituted
        back into \cref{eq:node} in order to eliminate $y$. This gives us a new
        equation, namely
        \begin{equation}
            \notag
            g(x, v) = x^2(v^2 - 1 - x).
        \end{equation}
        This function $g(x, v)$ is what we call an \emph{affine
        chart}\index{affine chart} of the blowup-surface $\Bl_p\A^2$. It can be
        visualized in the $(x, v)$-plane.

@@

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y).
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.

@@

        \begin{figure}[h]
            \centering
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure}
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly
                two points. This corresponds to the double point singularity in the
                original surface.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

@@

    In this section we will consider the algebraic surface given by the
    polynomial 
    \begin{equation}
        \label{eq:whitney}
        f(x, y, z) = x^2y - z^2.
    \end{equation}
    The associated variety is $\V(f(x, y, z))$. This surface is commonly known
    as the ''Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
    \cref{fig:whitney_umbrella}. As we will see shortly, when trying to blow up
    this surface in a single point, say the origin, the resulting surface does
    not seem to be any ''less singular''.

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
        \caption{The ''Whitney Umbrella'' defined by the polynomial $f(x, y, z)
        = x^2y - z^2$. This surface exhibits singularities all along the
        $y$-axis, as can easily be pointed out on the picture.}
        \label{fig:whitney_umbrella}
    \end{figure}

@@

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_1.png}
                \caption{The surface given by the equation $x^2(y - t^2) = 0$. The
                    curve of intersection between this surface and the exceptional
                    divisor looks like it can be ''pinched'' into a double line, i.e., the
                original line of singularity.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_2.png} 
                \caption{The surface given by the equation $r^2(z^2y - 1) = 0$. The
                    intersection between this surface and the exceptional divisor can
                    also be ''fused'' together to form a double line, by laying the
                surfaces flat on top of each other.}
            \end{subfigure}
            \\
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_1_exceptional.pdf} 
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_2_exceptional.pdf} 
            \end{subfigure}
            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The top
                row is rendered by \textsc{Surfer}, and the bottom row rendered by
                \textsc{Mathematica.} The exceptional divisor in the top row has
                been divided out and is therefore not shown because \textsc{Surfer}
                does a poor job at rendering planes. As seen, the charts are smooth and
                exhibits no singularities. The intersection with the exceptional divisor is
            rendered in blue in the bottom row.}
        \end{figure}
    \end{example}

    While in the above example we blew up a very specific line of
    singularities, namely the $y$-axis, it can be of interest to talk a little
    bit about how to blow up general lines. It all boils down to generalizing
    the construction of the plane for a more general line $L$. Assuming $L$ is
    parametrized by $x = at, y = bt, z = ct$, with $a, b$ and $c$ non-zero, we
    can form the symmetric equations 
    \begin{equation}
        \notag
        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
    \end{equation}
    The line $L$ is then completely determined by two of the three possible
    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
    the respective plane, we are interested in knowing when the plane $rx + sy
    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
    \begin{equation}
        \notag
        A = \begin{bmatrix}
            b & 0 & r \\
            -a & c & s \\
            0 & -b & t
        \end{bmatrix}
    \end{equation}
    one of the columns of $A$ is linearly dependent of the two others if
    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
    \begin{equation}
        \notag
        \det(A) = b(ct + sb) + rab = 0.
    \end{equation}
    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
    our plane.  The plane is then determined by 
    \begin{equation}
        \notag
        \frac{-ct - sb}{a}x + sy + tz = 0.
    \end{equation}
    Multiplying through by $a$ we achieve
    \begin{align*}
        (-sb - ct)x + sya + tza &= -sbx - ctx + sya + tza \\
                                &= t(za - cx) + s(ya - bx) = 0,
    \end{align*}
    which gives us the restrictions that we need.
    \clearpage
    \printbibliography \printindex
\end{document}

>>>>

\input{code_style.tex}
\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx} 
\makeindex

@@

\usepackage[noabbrev]{cleveref}

@@

    In algebraic geometry the central objects of study is the set of solutions
    to polynomial equations. Finding the zeros of, as an example, the quadratic
    equation
    \begin{equation}
        \notag
        ax^2 + bx + c = 0
    \end{equation}
    is something most people that have been exposed to elementary mathematics
    have done. Later on, one learns how to find the zeroes of more general
    functions, of either one or several variables over some general field of
    coefficients. Of special interest to us, are the polynomials over one or
    several variables, given as elements of a \emph{polynomial ring} over some
    field which we denote $K[x_1, x_2, \ldots, x_n]$.

    As an example, the most familiar polynomial ring is $\R[x]$. Elements of
    this ring include
    \begin{align*}
        f_1(x) = x^2, && f_2(x) = \pi, && f_3(x) = x^{17} + x^{3} - e^{4}, && f_4(x) = x^2 + 1.
    \end{align*} 
    Note that some of these have zeroes in the real numbers $\R$ while some
    does not. Most notably is the function $f_4$ whose zeroes lie in the realm
    of complex numbers, $\C$.

    Given a set of polynomials, $f_1, f_2, \ldots, f_m$ in a polynomial ring
    $K[x_1, x_2, \ldots, x_n]$ we can look at their common zeroes. It is in our
    interest to restrict our attention to polynomial rings where the field of
    coefficients is algebraically closed.

    Given an algebraically closed field $K$, we define \emph{affine
    $n$-space}\index{affine $n$-space} over $K$ to be the set of all $n$-tuples
    of elements of $K$. We denote affine $n$-space by $\A_K^n$ or $\A^n$ if the
    field of coefficients is clear from context.  For our intents and purposes,
    we will be working over the algebraically closed field of complex numbers
    $\C$.

@@

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities. A singular point is, strictly speaking, a point on a surface
    where the behaviour of the surface is slightly strange, say the surface
    passes through itself. We are often interested in studying how a surface
    behaves locally near a point $p$. We do this using \emph{linearization at
    $p$}. This is what we do when we consider the tangent line to a point for a
    single-valued function, the tangent plane for a general surface in three
    dimensions, and a tangent \emph{hyperplane} for general surfaces in $n$
    dimensions. The linearization of a surface given by $f(x_1, x_2, \ldots,
    x_n) = 0$, at a point $p_0$ is given by
    \begin{equation}
        \notag
        L(p) = f(p_0) + J_f(p_0)(p - p_0)
    \end{equation}
    where $J_f(p_0)$ is the Jacobian matrix of $f(x_1, x_2, \ldots, x_n)$
    evaluated at $p_0$. If this matrix has rank less than $n$, then the
    linearization is not uniquely defined, i.e., there are more than one plane
    satisfying the linearization. This leads us to the next defintion.

    \begin{definition}[Singularity {\cite[p.~3]{Whi08}}\index{singularity}]
        A singularity of the variety $V = \V(f_1, f_2, \ldots, f_m) \subset
        \A_K^n$ is a point $p \in V$ such that the Jacobian matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f_1}{\partial x_1} & \hdots & \frac{\partial f_1}{\partial x_n} \\
                \vdots & \ddots & \vdots \\
                \frac{\partial f_m}{\partial x_1} & \hdots & \frac{\partial f_m}{\partial x_n}
            \end{bmatrix}(p)
        \end{equation}
        has rank strictly less than $\min(m, n)$.
    \end{definition}

    A very useful consequence of this definition deal with the varieties of
    single polynomials and will aid us tremendously in the search for
    singularities.
    \begin{corollary}
        \label{crl:single_poly}
        A point $p$ is a singularity of the variety $V = \V(f) \subset \A_K^n$
        if and only if the partial derivatives of $f$ vanish identically at
        $p$. 
    \end{corollary}
    \begin{proof}
        We recall that the rank of a matrix is the dimension of its column
        space, that is for a matrix $A$, $\rank(A) = \dim(\Col(A))$.  Assume
        first that $p$ is a singularity of $V$. Then $\rank(A) < min(n, m) =
        1$. Consequently, we must have $\rank(A) = 0$, and therefore $A$ equal
        to the zero matrix. All the partial derivatives are therefore zero.
        For the opposite implication, assume that the partials vanish
        identically at $p$. Then the Jacobian matrix reads
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                0 & \hdots & 0
            \end{bmatrix},
        \end{equation}
        and consequently $A$ has rank strictly less than one. Hence $p$ is a
        point of singularity of $V$.
    \end{proof}

    Finally, we want to define the notion of a projective space, as this will
    aid us in the resolution of singularities later on.  Informally, projective
    space generalizes the notion of parallel lines intersecting at infinity.
    The general definition of projective $n$-space is the set of lines in
    $\A^{n+1}$ passing through the origin. For our purposes we will take
    $\A^{n+1} = \C^{n+1}$.

@@

    \section{Introduction to algebraic surfaces}

@@

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ``The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ``The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{Curves not corresponding to functions.}
        \label{fig:algebraic_curves}
    \end{figure}

@@

    From \cite[p.~7]{Lab14} we have the following upper bound on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(d) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

@@

    We call a variety that exhibits no singularities \emph{smooth}\index{smooth
    variety}. It is often possible to study a smooth copy of a variety with
    singularities if we move from affine space to projective space. This
    process can be defined in many different ways, with the seemingly most
    prevalent one being through maps from various spaces to each other. In this
    paper we will try to keep it as intuitive as possible, by using the more
    geometrically based procedure.

@@

    We start by examining resolutions of single points. Informally, when we
    blow up a point $p$ in $\A^n$, we replace the point $p$ by an entire copy
    of a projective space $\P^{n-1}$ and try to leave the rest of the space
    $\A^n$ ``unchanged''. We will give a few examples, that hopefully will
    illustrate the geometrical aspect of this, but first a formal definition.

    \begin{definition}[Blowup of $\A^{n}$ at a point $p$.]
        We let the \emph{blowup-surface}\index{blowup-surface!in a point}
        $\Bl_p\A^{n}$ be the set of all pairs consisting of a point $p \in
        \A^n$ and a line from $p$ through the origin. The surface is embedded
        in the space $\A^n\times\P^{n-1}$. We write
        \begin{equation}
            \notag
            \Bl_p\A^n = \left\{ (p, L) \mid p \in \A^{n} \text{ and } p \in L \right\}.
        \end{equation}
    \end{definition}
    Now, this is a fairly technical definition, so where do we go from here?
    When blowing up a surface at a point $p$ we are essentially giving the
    surface ``more room'' so that the points creating the singularities can
    pass each other without hassle. In order to aid us with the above
    definition, we show the following claim that tells us under what
    circumstances the point $p$ lie on the line $L$:

    \begin{proposition}
        \label{prop:point_line}
        Given a point $x = (x_1, x_2, \ldots, x_n)$ in $\A^n$ and a point $y =
        [y_1 : y_2 : \ldots : y_n]$ in the space $\P^{n-1}$, the point $x$ is a
        multiple of $y$ if and only if the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x_1 & x_2 & \hdots & x_n \\
                y_1 & y_2 & \hdots & y_n
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to 1. Similarly, the rank is less than or
        equal to 1 if and only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
        $1 \leq i < j \leq n$.
    \end{proposition} 
    \begin{proof}
        In order to verify this, we recall that $\rank(A) = \rank(A^T)$, so we
        chose to work with the transposed matrix instead. Assume that $x$ is a
        multiple of $y$, so $y = \lambda x$ for some scalar $\lambda$. Then
        \begin{equation}
            \notag
            A^T = \begin{bmatrix}
                x_1 \;\, \lambda x_1 \\
                x_2 \;\, \lambda x_2 \\
                \vdots \\
                x_n \;\, \lambda x_n
            \end{bmatrix}
        \end{equation}
        is the associated matrix. If $x = 0$, then $A^T$ is the zero matrix and
        has therefore rank 0. If $x \neq 0$, then we have exactly one pivot
        column in this matrix, hence $\dim(\Col(A^T)) = 1$. In any case,
        $\rank(A) = \rank(A^T) \leq 1$. For the opposite implication. If $A$
        has rank 0, then $A$ is the zero matrix, and hence trivially, $x$ is a
        multiple of $y$, both being zero. If $A$ has rank $1$, then there is
        one linearly independent column in $A^T$, hence the other column can be
        expressed as a multiple of the first. In other words, $x$ is a multiple
        of $y$.

@@

        Setting $u = 1$ we get the equation $y = xv$. This can be substituted
        back into \cref{eq:node} in order to eliminate $y$. This gives us a new
        equation, namely
        \begin{equation}
            \notag
            g(x, v) = x^2(v^2 - 1 - x) = 0.
        \end{equation}
        The variety $\V(g(x, v))$ is what we call an \emph{affine
        chart}\index{affine chart} of the blowup-surface $\Bl_p\A^2$. It can be
        visualized in the $(x, v)$-plane.

@@

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y) = 0.
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.

@@

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure}
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly
                two points. This corresponds to the double point singularity in the
                original surface.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

@@

    In this section we will consider the algebraic surface given by the
    polynomial 
    \begin{equation}
        \label{eq:whitney}
        f(x, y, z) = x^2y - z^2.
    \end{equation}
    The associated variety is $\V(f(x, y, z))$. This surface is commonly known
    as the ``Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
    \cref{fig:whitney_umbrella}. As we will see shortly, when trying to blow up
    this surface in a single point, say the origin, the resulting surface does
    not seem to be any ``less singular''.

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
        \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
        = x^2y - z^2$. This surface exhibits singularities all along the
        $y$-axis, as can easily be pointed out on the picture.}
        \label{fig:whitney_umbrella}
    \end{figure}

@@

        The two affine charts are shown visualized in \cref{fig:affine_line}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_1.png}
                \caption{The surface given by the equation $x^2(y - t^2) = 0$. The
                    curve of intersection between this surface and the exceptional
                    divisor looks like it can be ``pinched'' into a double line, i.e., the
                original line of singularity.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_2.png} 
                \caption{The surface given by the equation $r^2(z^2y - 1) = 0$. The
                    intersection between this surface and the exceptional divisor can
                    also be ``fused'' together to form a double line, by laying the
                surfaces flat on top of each other.}
            \end{subfigure}
            \\
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_1_exceptional.pdf} 
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/affine_2_exceptional.pdf} 
            \end{subfigure}
            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The top
                row is rendered by \textsc{Surfer}, and the bottom row rendered by
                \textsc{Mathematica.} The exceptional divisor in the top row has
                been divided out and is therefore not shown because \textsc{Surfer}
                does a poor job at rendering planes. As seen, the charts are smooth and
                exhibits no singularities. The intersection with the exceptional divisor is
            rendered in blue in the bottom row.}
            \label{fig:affine_line}
        \end{figure}
    \end{example}

    While in the above example we blew up a very specific line of
    singularities, namely the $y$-axis, it can be of interest to talk a little
    bit about how to blow up general lines. It all boils down to generalizing
    the construction of the plane for a more general line $L$. Assuming $L$ is
    parametrized by $x = at, y = bt, z = ct$, with $a, b$ and $c$ non-zero, we
    can form the symmetric equations 
    \begin{equation}
        \notag
        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
    \end{equation}
    The line $L$ is then completely determined by two of the three possible
    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
    the respective plane, we are interested in knowing when the plane $rx + sy
    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
    \begin{equation}
        \notag
        A = \begin{bmatrix}
            b & 0 & r \\
            -a & c & s \\
            0 & -b & t
        \end{bmatrix}
    \end{equation}
    one of the columns of $A$ is linearly dependent of the two others if
    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
    \begin{equation}
        \notag
        \det(A) = b(ct + sb) + rab = 0.
    \end{equation}
    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
    our plane.  The plane is then determined by 
    \begin{equation}
        \notag
        \frac{-ct - sb}{a}x + sy + tz = 0.
    \end{equation}
    Multiplying through by $a$ we achieve
    \begin{align*}
        (-sb - ct)x + sya + tza &= -sbx - ctx + sya + tza \\
                                &= t(za - cx) + s(ya - bx) = 0,
    \end{align*}
    which gives us the restrictions that we need.

    \clearpage 
    \appendix

    \section{Code snippets}
    \label{sec:code_snippets}

    \lstset{deletekeywords={xrange}}
    \begin{TMcode}{python}{curves.py - \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
    import matplotlib.pyplot
    from numpy import arange
    from numpy import meshgrid

    delta = 0.025
    xrange = arange(-5, 5.0, delta)
    yrange = arange(-2.5, 2.5, delta)
    X, Y = meshgrid(xrange,yrange)

    # F is one side of the equation, G is the other
    F = X**2*(1 - Y**2 - X*Y**3)
    G = 0

    matplotlib.pyplot.axis('on')
    matplotlib.pyplot.grid('on')
    matplotlib.pyplot.contour(X, Y, (F - G), [0])

    matplotlib.pyplot.savefig('output.pdf')
    \end{TMcode}

    \clearpage
    \begin{TMcode}{mathematica}{intersection.nb - \url{http://community.wolfram.com/groups/-/m/t/178218?p_p_auth=lT7er3Rd}}{Plotting the intersection between to surfaces}

    (* affine chart and exceptional divisor *)
    h = z^2 y - 1;
    g = x^2;
